Parameter precision in economic growth models inspired by Ecological Economics

### 2015-10-05 Articles/Article_EcolEcon_2015/Paper0/Paper0-Resampling.Rnw
Editing methods section.

The Constant Elasticity of Substitution~(CES) model 
is given as
%
\begin{equation} \label{eq:CESgeneric}
  y = \theta \: A \: \left\{\delta B_1^{\rho/\rho_1} 
      + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda t},
\end{equation}
%
where
$B_1 \equiv \delta_1 x_1^{-\rho_1} + (1-\delta_1) x_2^{-\rho_1}$, and
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of capital~($k$), labor~($l$), and energy~($e$).

@@

\begin{equation} \label{eq:alpha_2}
  \alpha_2 \equiv \frac{ \frac{\partial y}{\partial x_2} }{\frac{y}{x_2}} =
  \frac{ \delta  \left(1-\delta _1 \right) x_2^{-\rho _1} B_1^{ ( \frac{\rho }{\rho _1}-1 ) } }
  { \delta B_1^{ ( \frac{\rho }{\rho_1} ) } + (1 - \delta) x_3^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_3}
  \alpha_3 \equiv \frac{ \frac{\partial y}{\partial x_3} }{ \frac{y}{x_3} } =  
      \frac{1 - \delta}{\delta  x_3^{\rho } B_1^{(\frac{\rho }{\rho_1} ) } + 1 - \delta}
\end{equation}
%
It can be verified that
%
\begin{equation} \label{eq:sum_betas}
  \alpha_1 + \alpha_2 + \alpha_3 = 1
\end{equation}
%
for all nests.

@@

To estimate values of parameters
($\theta$, $\lambda$, $\delta_1$, $\delta$, $\rho_1$, and $\rho$)
for the CES model,
we use an ordinary least squares~(OLS) approach.
The objective of the OLS analysis is minimization of the sum of squared errors~(Equation~\ref{eq:sse}).
We estimate parameters using the \texttt{R}~\citep{R} package \texttt{micEconCES}
\citep{Henningsen:2011td}
with significant customization
to estimate parameters in the CES production function.
Our CES parameter estimation algorithm starts with an eleven-value grid search
in $\rho_1$ and $\rho$
(9, 2, 1, 0.43, 0.25, 0.1, -0.1, -0.5, -0.75, -0.9, -0.99),
corresponding to $\sigma_1$ and $\sigma$ values of
0.1, 0.33, 0.5, 0.7, 0.8, 0.9, 1.11, 2, 4, 10, and 100,
respectively.
During the grid search, values of $\rho_1$ and $\rho$ are fixed,
and values of $\theta$, $\lambda$, $\delta_1$, and $\delta$ are estimated
by gradient search with the \texttt{PORT} and \texttt{L-BFGS-B} algorithms
in the \texttt{micEconCES} package~\citep{Henningsen:2011td}.
In all, 121 gradient searches in $\theta$, $\lambda$, $\delta_1$, and $\delta$
at grid points representing all combinations of $\rho_1$ and $\rho$ are
attempted.
During the grid search portion of our algorithm, starting values for the free
parameters
are $\lambda$ = 0.015/year, $\delta_1$ = 0.5, $\delta$ = 0.5,
and $\theta$ is set by the \texttt{cesEst} function
to a value such that the mean of the residuals is zero.

Next, a gradient search (using both \texttt{PORT} and \texttt{L-BFGS-B})
is attempted wherein all parameters
($\theta$, $\lambda$, $\delta_1$, $\delta$, $\rho_1$, and $\rho$)
are allowed to float.
The starting values for fitting parameters are taken from
the grid search point that provided the lowest sum of squared errors~(\sse).

@@

In the context of linear models (regression),
resamples are generally created
by residual resampling.
In our case, we formed resamples by adding to the fitted
response~($\hat{y}$) the product of a residual from the original CES function fit
(see Figure~\ref{fig:r_kle}) and
random sign ($-1$ or $1$, each with probability $0.5$).
This residual resampling occurs on the
log-transformed data.
Intuitively, this method assumes that the residuals are indicative of
the variability (from many potential sources) inherent in the data such that
it would be unsurprising if the residual from
any particular year had been observed in a different year.
Thus, a resampled response $\tilde Y'$ can be computed as%
\authNote{Randy: edit the following equations to reflect the fact that
we're fitting in log space? ---MKH}

\begin{equation} \label{eq:resampled_y}
  \tilde Y_i' = \hat y_i \pm r_j \; ,
\end{equation}
%
where
%
\authNote{Do we need this equation for $r_i$ here? We already have Equation~\ref{eq:residual}.}%
%
\begin{equation} \label{eq:residual_2}
  r_i = y_i - \hat y_i \; .
\end{equation}
%
Both the sign~($\pm$) and the index
($j$,~typically different from $i$)
of the residual~($r_j$)
are chosen at random (with replacement).
We repeated the resampling process 1000 times for each
combination of growth model and country,
both with and without energy as appropriate.

@@

We choose the resampling approach instead of the more-common technique
of estimating confidence intervals from standard errors for two reasons.
First, model parameters are highly constrained, and confidence intervals often violate the constraints.
For example, a result of $\alpha_k = 0.25 \pm 0.3$ is nonsensical,
because $ \alpha_k \in [0, 1]$.
Second, true confidence intervals for constrained parameters are often asymmetric
(e.g., $\alpha_k = 0.2 + 0.1, -0.05$),
but the standard error approach yields symmetric confidence intervals.
The resampling approach that we employed both respects constraints
and allows for asymmetric confidence intervals.%
In addition, we choose to present parameter uncertainty results visually
(see Figures~\ref{fig:resample_lambda_theta}--\ref{fig:resample_alpha}),
both because the resampling technique lends itself to visual presentation and
because we believe visual comparison is more effective
in communicating results than endless tables of numbers.

>>>>

The Constant Elasticity of Substitution~(CES) model 
is given as
%
\begin{equation} \label{eq:CESgeneric}
  y = \theta \: A \: \left\{\delta B_1^{\rho/\rho_1} 
      + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda t},
\end{equation}
%
where
$B_1 \equiv \delta_1 x_1^{-\rho_1} + (1-\delta_1) x_2^{-\rho_1}$, and
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of capital~($k$), labor~($l$), and energy~($e$).
The $(kl)e$ nesting has $x_1 = k$, $x_2 = l$, and $x_3 = e$; 
the $(le)k$ nesting has $x_1 = l$, $x_2 = e$, and $x_3 = k$; and
the $(ek)l$ nesting has $x_1 = e$, $x_2 = k$, and $x_3 = l$.

@@

\begin{equation} \label{eq:alpha_2}
  \alpha_2 \equiv \frac{ \frac{\partial y}{\partial x_2} }{\frac{y}{x_2}} =
  \frac{ \delta  \left(1-\delta _1 \right) x_2^{-\rho _1} B_1^{ ( \frac{\rho }{\rho _1}-1 ) } }
  { \delta B_1^{ ( \frac{\rho }{\rho_1} ) } + (1 - \delta) x_3^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_3}
  \alpha_3 \equiv \frac{ \frac{\partial y}{\partial x_3} }{ \frac{y}{x_3} } =  
      \frac{1 - \delta}{\delta  x_3^{\rho } B_1^{(\frac{\rho }{\rho_1} ) } + 1 - \delta}
\end{equation}
%
It can be verified that
%
\begin{equation} \label{eq:sum_betas}
  \alpha_1 + \alpha_2 + \alpha_3 = 1 \; .
\end{equation}

@@

To estimate values of parameters
($\theta$, $\lambda$, $\delta_1$, $\delta$, $\rho_1$, and $\rho$)
for the CES model,
we use an ordinary least squares~(OLS) approach, 
wherein the objective is minimization of the sum of squared errors~(Equation~\ref{eq:sse}).
We estimate parameters using the \texttt{R}~\citep{R} package \texttt{micEconCES}
\citep{Henningsen:2011td}
with significant customization
to estimate parameters.
Our CES parameter estimation algorithm starts with an eleven-value grid search
in $\rho_1$ and $\rho$
(9, 2, 1, 0.43, 0.25, 0.1, -0.1, -0.5, -0.75, -0.9, -0.99).
During the grid search, values of $\rho_1$ and $\rho$ are fixed,
and values of $\theta$, $\lambda$, $\delta_1$, and $\delta$ are estimated
by gradient search with the \texttt{PORT} and \texttt{L-BFGS-B} algorithms
in the \texttt{micEconCES} package~\citep{Henningsen:2011td}.
In all, 121 gradient searches in $\theta$, $\lambda$, $\delta_1$, and $\delta$
at grid points representing all combinations of $\rho_1$ and $\rho$ are
attempted.
During the grid search portion of our algorithm, starting values for the free
parameters
are $\lambda$ = 0.015/year, $\delta_1$ = 0.5, $\delta$ = 0.5,
and $\theta$ is set by the \texttt{cesEst} function
to a value such that the mean of the residuals is zero.

Next, a gradient search (using \texttt{PORT} and \texttt{L-BFGS-B} algorithms available in \texttt{micEconCES})
is attempted wherein all parameters
($\theta$, $\lambda$, $\delta_1$, $\delta$, $\rho_1$, and $\rho$)
are allowed to float.
The starting values for fitting parameters are taken from
the grid search point that provided the lowest sum of squared errors~(\sse).

@@

In the context of linear models (regression),
resamples are generally created
by residual resampling.
In our case, we formed resamples by adding to the fitted
response~($\hat{y}$) the product of a residual from the original function (Cobb-Douglas or CES) fit
and
random sign ($-1$ or $1$, each with probability $0.5$).
This residual resampling occurs on the
log-transformed data.
Intuitively, this method assumes that the residuals are indicative of
the variability (from many potential sources) inherent in the data such that
it would be unsurprising if the residual from
any particular year had been observed in a different year.
Thus, a resampled response $\tilde Y'$ can be computed as%
\authNote{Randy: edit the following equations to reflect the fact that
we're fitting in log space? ---MKH}

\begin{equation} \label{eq:resampled_y}
  \tilde Y_i' = \hat y_i \pm r_j \; ,
\end{equation}
%
where
%
\authNote{Do we need this equation for $r_i$ here? We already have Equation~\ref{eq:residual}.}%
%
\begin{equation} \label{eq:residual_2}
  r_i = y_i - \hat y_i \; .
\end{equation}
%
Both the sign~($\pm$) and the index
($j$,~typically different from $i$)
of the residual~($r_j$)
are chosen at random (with replacement).
We repeated the resampling process 1000 times.

@@

We choose the resampling approach instead of the more-common technique
of estimating confidence intervals from standard errors for two reasons.
First, model parameters are highly constrained, and confidence intervals often violate the constraints.
For example, a result of $\alpha_1 = 0.25 \pm 0.3$ is nonsensical,
because $ \alpha_1 \in [0, 1]$.
Second, true confidence intervals for constrained parameters are often asymmetric
(e.g., $\alpha_1 = 0.2 + 0.1, -0.05$),
but the standard error approach yields symmetric confidence intervals.
The resampling approach that we employed both respects constraints
and allows for asymmetric confidence intervals.
